The importance of measurement uncertainty in ecological management

### 2017-12-13 manuscripts/manuscript.Rmd
separate out old / scratch drafts from current working draft

Imperfect information is ubiquitous in ecological management and conservation decision making. While the pressing concerns of global change have put the spotlight on forecasting, i.e. the uncertainty of the future [@Petchy2015], management decisions must also contend with uncertainty of the present: How many fish are in the sea today? What regions harbor the most fragile biodiversity? How have invasive species transformed native communities? A recognition of the importance of uncertainty as well as sophisticated methods and tools to do so has become increasingly widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierachical Bayesian Modeling, along with convenient software tools for implementing these methods.  In sharp contrast to that shift, the treatment of uncertainty in *decisions* -- how the uncertainty in models should be reflected in scientific recommendations, remains far less developed, more qualitative, and underserved by available tools.  In this paper we illustrate how uncertainty in measurement has a dramatic negative impact on outcomes under a variety of these exsiting approaches, which can be avoided through the use of more formal approaches that have since been developed in other fields that can likewise be implemented in software tools.

@@

The area of fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management, including topics such as adaptive management [@Walters], ecosystem-based management [@Levin2008] to resilience thinking [@Holling; @May] and back to the birth of resource economics [@Beverton; @Scheffer], thanks to its global relevance, long history, and readily available data.  Consequently, understanding the impact of measurement uncertainty in this rich and well-studied context will have implications for both resource management and conservation efforts in other domains, many of which draw on concepts established or tested in fisheries ecosystems.  We compare the results of managing simulated fish stocks under the existing optimal control approach, a heuristic approach that more closely resembles today's management practices, and a newly proposed approach that provides a more rigorous treatment of uncertainty.

@@

Numerically, each of these functions are defined over a discrete set of possible states, observations, and actions, and can thus be represented as a collection of matrices or tensors.  As in the case of MDP [@Marescot2013], the POMDP problem can be solved by existing and well-studied algorithms. While not unknown to the conservation literature [@Williams2009], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g. @Chades, @Fackler, ....].  POMDP algorithms remain a very active area in artificial intellegence, and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries managment, which requires on the order of 100 states and actions to provide sufficient numerical resolution. Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods. For those interested in exploring or applying these methods, we have developed a preliminary R package, @sarsop, which can be used to replicate any of the results presented here and explore other variations.

@@

@@

-----------------
<!-- MISC draft text -->

It is precisely this reasoning which explains the broad patterns of recovery
observed in the landmark paper @Costello2016, which compares future of global
fisheries under a Rights Based Fisheries Mangement (RBFM), fishing at MSY,
and "buisness as usual" (BAU) scenario.  Their analysis assumes that a RBFM
approach should achieve the economically optimal policy of constant-escapement,
which alone explains why RBFM should see both faster recovery and higher 
economic yields than MSY, as we have just described. The substantial detail their
analysis introduces with respect to the dynamics of over 4000 individual stocks
as well as additional economic knock-on effects from managing based on quotas
rather than fishing effort permit them to offer quantitatively precise 
estimates regarding the time required and economic value generated by such
recovery, but are not essential to understand the broad pattern that suggests
that nearly all stocks should recover under either an RBFM or MSY-style 
management, but that the recovery would be fastest and the economic benefits
greatest under the former.  That conclusion was already implicit in how
these models are defined.

-------------

A fundamental concept to the management of any natural population is the notion of Maximum Sustainable Yield, MSY. The notion of MSY is rooted in the determinstic models dating back to foundational work in resource economics and ecological management half a century ago [@Berverton; @Scheffer].  Though the concept has been the focus of significant criticism for some time [e.g. @Botsford_nature_piece] that ctitique has been largely on the "Maximum" part, while the existence of a "Sustainable Yield."

## Reed's Paradox: $S = D$

In 1979, a fundamental paper by William Reed [@Reed1979] proved that, under a very general assumptions, the harvest policy for a stochastically fluctuating population was identical to that of the equivalent, deterministic population.  More precisely, Reed shows that the optimal harvest of the stochastically flutuation population should always aim to reduce the population back to a constant size $S$ (termed by Reed as "escapement," while in most fisheries literature uses the term $B_{MSY}$, the biomass at Maximum Sustainable Yield}).  Moreover, this value, $S$, is no more conservative than had we been able to predict future populations with perfect accuracy, as in the determinstic case of constant escapement, $D$.  A notable exception to this result is when the population is not "self-sustaining," that is,
the stochasticity is severe enough to pose an extinction risk to the unhavested population.  In such cases Reed proves there is still a single constant target population size $S$, but that $S > D$, though often by very small amount, reflecting the small probability of a purely chance extinction.

An important corrolary of this result is that suggested a way to side-step the laborious dynamic programming required to find an optimal policy for stochastic populations: the deterministic solution is found trivially from maximum of the stock-recruitment curve. Thus much subsequent attention has focused on the estimating stock-recruitment, including methods that address considerable uncertainty in the data used for this estimate [@], while largely ignoring the subsequent step of how to determine an optimal policy in light of that uncertainty.

The lack of any intuitively precautionary response to the introduction of uncertainty did not go unnoticed: in 1986 Clark and Kirkwood first pointed to the absence of measurement error in the formulation used by Reed and others as fundamental to understanding this paradox.

Under such an assumption, uncertainty introduced by stochastic growth is effectively purged by each subsequent observation rather than accumulating, and stock extinction can only occur if the manager explicitly decides to harvest the entire stock.

## Clark's Paradox: M < D

Clark and Kirkwood set out to address this problem by considering the case in which stochastic population growth occurs after population size is measured but before harvest, such that manager must act without perfect knowledge of the current system state.  Unfortunately, this problem proved easier to recognize than to solve, as Clark & Kirkwood recognized at the outset:

Rather than resolve Reed's Paradox, this assumption compounds the problem, leading to optimal harvests that are generally even higher (i.e. less cautious) than the analgous deterministic policy.  The same counter-intuitive result can also be found in subsequent work [@Sethi2005] which makes a different but equally limiting simplifying assumption (supplement).  (Both @Clark1986 and @Sethi2005 remark on the counter-intuitive nature of their results)

<!--

Clark: 
> The above results appear to contradict the conventional wisdom of renewable resource management, under which high uncertainty would call for increased
caution in the setting of quotas.

Sethi: 
> It may seem counter-intuitive that a measurement error causes lower expected escapements below the deterministic fishery closure threshold.

-->

## Precautionary approaches

Perhaps recognizing

## A solution

Here, we explore how uncertainty over the present impacts decisions

We shall focus on examples from fisheries, which have long been both incubator and proving ground for most of our more general theories of ecological management, from the and

Data are readily avialable and a compartively strong connection between scientific process of stock assessment and the implementation of policy is already established.  It is also an area which, as we will see, though many models already account for uncertainty, decision strategy still reflects fundamentally more deterministic ideas.

Conclusions:

- failing to account for uncertainty properly can lead to erroneous conclusions
- Prior beliefs (reflecting prior observations) matter.

<!-- FIXME: In practice, MSY and Reed are very different -->

<!-- Big picture: fisheries is a foundation and testing ground for conservation
and management. Yet at it's heart lies a controdiction.  -->

<!-- MSY and Constant Escapement -->

Yet fisheries are not managed at their maximum ecological or economic potential.  
Taken as a whole, most ocean fisheries yield a net negative economic return
[@FAO2009], while some 68% of global fisheries are not in good biological condition [@Costello2016].
Nevertheless, fisheries management in both theory and practice are

In a deterministic world at equilibrium (without economic discounting), 
the policies of constant escapement and maximum
sustainable yield look identical: the stock is maintained at a fixed stock
size or biomass, called $B_{MSY}$ in the MSY/fisheries jargon, which is 
equivalent to the escapement $S^*$ in the optimal control jargon and equal
to half the carrying capacity, or $K/2$, in the familiar Graham-Schaefer 
(logistic growth followed by harvest) model. Likewise, the harvest is identical,
occuring after recruitment and bringing the population back to this optimal
level; an annual harvest of $H_{MSY} = rK/4$ in Graham-Schafer model for either policy.
This similarity disguises significant differences in how the policies behave
when the biomass is either above or below $B_{MSY}$.

<!-- Ways in which constant escapement deviates from MSY -->
The economically optimal amount to harvest when the population is below 
$B_{MSY}$ is zero -- no harvest.  This permits the population to return
as rapidly as possible to the stock size at which it generates the maximum
yield, and thus the maximum economic return.  This result holds even when
recruitment is stochastic and thus future stock sizes are uncertain, 
as rigorously proven by @Reed1979.  Perhaps surprisingly, this is not what
fisheries management thinks of as a maximum sustainable yield (MSY) policy,
and does not reflect how fisheries are managed today. Rather, an MSY policy
is intepreted as fishing at the *effort*, $F_{MSY} = H_{MSY} / B_{MSY}$ that would generate the maximum
sustainable yield if the population was at the optimal biomass, $B_{MSY}$,
*regardless* of the current stock size. It is important to realize this policy
is framed in terms of *effort* $F$: for example, the length of the year open to fishing,
or equivalently, the per-capita mortality rate of fish, and not in terms of harvest, $H$.
A population below $B_{MSY}$ that remains harvested at a constant $H_{MSY}$ will eventually be
driven to extinction, but harvested at a constant $H = F_{MSY} \cdot B$ for $B < B_{MSY}$,
the stock will eventually recover to $B_{MSY}$.  Consequently, this simple policy should
converge to the optimal policy in the long run, assuming there is no uncertainty in
recruitment, measurement, or implementation.

The economically optimal policy of constant-escapement also differs from MSY
when stocks are above $B_{MSY}$.  As before, the MSY policy continues to assume
a fixed mortality rate, $F_{MSY}$, which should eventually reduce the stock to
the level of $B_{MSY}$.  Also as before, the optimal policy is not so patient --
to derive maximum yield it seeks to return the population to the $B_{MSY}$ not
*eventually*, but as quickly as possible, and thus will recommend a harvest larger
than $H_{MSY}$ to accomplish this. These differences become more salient in the case
of a population under stochastic growth, since the resulting variation leads
to a population that frequently above or below $B_{MSY}$.  The greater responsiveness
of the constant-escapement policy to the current stock size thus necessarily 
results in both the faster recovery of overexploited stocks and higher economic
returns.

The variation in the harvest required to achieve this constant escapement policy
can seem impractical from a top-down management perspective.  However, it is 
important to realize that the application and implications of this policy are
not purely academic. Economists have for some time observed these inherent 
inefficiencies, and argued that the establishment of property rights in fisheries [@Arnason2012].
Approaches such as individual transferable quotas (ITQs) would incentivise
fisheries behavior to converge towards the more dynamic harvests of a constant
escapement policy, as doing so would increase the value of an individual's share.
It is precisely this reasoning which explains the broad patterns of recovery
observed in the landmark paper @Costello2016, which compares future of global
fisheries under a Rights Based Fisheries Mangement (RBFM), fishing at MSY,
and "buisness as usual" (BAU) scenario.  Their analysis assumes that a RBFM
approach should achieve the economically optimal policy of constant-escapement,
which alone explains why RBFM should see both faster recovery and higher 
economic yields than MSY, as we have just described. The substantial detail their
analysis introduces with respect to the dynamics of over 4000 individual stocks
as well as additional economic knock-on effects from managing based on quotas
rather than fishing effort permit them to offer quantitatively precise 
estimates regarding the time required and economic value generated by such
recovery, but are not essential to understand the broad pattern that suggests
that nearly all stocks should recover under either an RBFM or MSY-style 
management, but that the recovery would be fastest and the economic benefits
greatest under the former.  That conclusion was already implicit in how
these models are defined.  Unfortunately,

-------

Measuring and understanding collapse and over-exploitation of fisheries
world-wide remains an essential issue for the ecological sciences
[@Costello2016].  Effective management practices in marine ecosystems also
inform conservation and resource management in less well-studied
systems.

<!-- Set up broad relevance of understanding this issue -->

Central to both fisheries research and
policy, including international agreements such as the United Nations
Law of the Sea [@UNCLOS1982], is the principle of  maximal sustainable yield (MSY).
MSY aims to maintain a population at a fixed population size which should result in
the greatest long-term harvest (often referred to as Biomass at MSY,
$B_{MSY}$). While the theory dates back to classic work
of 1950s such as Beverton [@Beverton1957] and Schaefer [@Schaefer1957]
in deterministic models, it is Reed 1979 paper [@Reed1979] which proves
that the intuitive result of the deterministic models remains the optimal
policy for maximizing expected value of a fishery subject to stochastic
growth (i.e. environmental noise). Reed's result has allowed management
to largely sidestep the complexity of optimal control of stochastic
systems.

This has long begged the question: if stochastic growth can be
safely ignored in determining the optimal harvest, is the same true of
measurement error?  Thanks to recent computational advances for solving
the decision theory problem in the case of imperfect observations
(measurement error), we show that -- contrary to both the intuition
from Reed's result and previous, approximate solutions -- ignoring
measurement error can result in significant over-harvesting, decreased
economic returns, and the potential for complete collapse of the stock.

Some historical perspective is necessary to put this result in the
appropriate context with respect to both management practice as well
as previous research in optimal control of fisheries and other natural
systems under uncertainty.  When population growth
varies stochastically, any fixed annual harvest such as MSY is no longer optimal:
instead, management must respond by dynamically harvesting more in years
of high growth and less in years of low growth.  Reed's proof [@Reed1979]
nevertheless demonstrated that the optimal harvest can still be captured
by a simple rule: harvesting just enough to reduce the population back
down to $B_{MSY}$.  This result is known as a "constant-escapement" policy 
because the population escaping harvest is always the same, 
even while the harvest itself varies dynamically with each stochastic shock. 
Moreover, Reed shows that under some relatively generic assumptions, this escapement
is identical to that predicted by the deterministic theory  ($B_{MSY}$).
Because the optimal harvest policy in the deterministic theory follows directly
from the growth rate (just find the stock size for which recruitment function
is maximized, no need for dynamic programming), attention shifted to inferring
that state equation from data.  Our focus is not on inferring this state
equation, but on the subsequent step of: having done so, how does one determine
the optimal harvest policy while also accounting for measurement uncertainty? 
Previous attempts to address this issue over the past three decades
have relied on various approximations which have usually led to the paradox
that harvests should increase under additional measurement uncertainty, contrary 
to our intuition that we should respond to greater uncertainty with greater caution.
Using more recent optimal control algorithms first developed in the artificial
intelligence literature, we provide a solution which avoids such assumptions
and resolves this paradox in favor of common sense intuition.

Despite it's generality, Reed's proof makes a subtle assumption which
reduces the the impact uncertainty and underlies his conclusion
that the deterministic intuition of an average MSY is still optimal in 
the stochastic case.
In Reed's construction, the manager measures the stock
size without error during each time step before setting the harvest policy.
An important consequence of this construction is that this model can
thus never lead to accidental extinction of the stock.
Researchers have long recognized the weakness of this assumption of
perfect knowledge of the stock size and have sought to understand how it
may impact the resulting management policy [@Ludwig1981].  Because there is no risk of
accidental extinction in this scenario, it is possible that the optimal
policy may be more aggressive than is warranted.  Unfortunately, due
to the complexity of the analysis required to account for measurement
uncertainty, this question has never been satisfactorily answered.
Moreover, previous attempts at an approximate solution have consistently
found the surprising conclusion that the Reed solution is usually less
aggressive (dictating higher escapement or lower catch quotas) than
attempts to account for measurement uncertainty.  We briefly review the
main results in the ecological and resource economics literature that
have attempted to address this through various approximations before
presenting our solution, which draws on substantial algorithmic advances
that have been made in other fields.

The important way in which perfect observations diminish the role of
uncertainty in the model of Reed [@Reed1979] was first recognized in the classic
work of Clark & Kirkwood [@Clark1986].  They acknowledge the importance of measurement
uncertainty to the optimal solution, while observing that if
this assumption was relaxed, the "difficulty of the problem increases
markedly." Instead, @Clark1986 consider a perfectly observed model in
which harvest takes place after stochastic recruitment but before the
stock is measured, making accidental extinction a possibility
should realized growth fall sufficiently short of expectation.
In this case, @Clark1986 demonstrate that the optimal solution deviates
from the constant escapement rule, with the resulting optimal
policy is often significantly less cautious than the constant escapement policy.
In certain cases, @Clark1986 demonstrate that harvesting to extinction
is actually optimal. Other recent work [@Holden2015; @Johnson2016] has taken 
a similar approach to @Clark1986 in which measurement becomes uncertain due to
stochastic stock dynamics occurring between measurement and harvest, which
can also suggest less conservative (higher harvest) escapement rates.

Rougharden [@Roughgarden1996] take a complementary approach, introducing measurement
uncertainty but searching only for the optimal constant escapement,
despite presenting no evidence (and indeed contrary to @Clark1986)
that the optimal solution under this circumstance should be a
constant escapement policy. By applying the rule-of-thumb assumption
that the optimal policy can be expressed simply as a constant escapement
(or equivalently, target-stock), the set of possible policies is reduced
dramatically and the calculation is relatively trivial.  As expected,
this restriction results in a policy that is always more conservative (lower harvest) than
the Reed solution which ignores measurement error, but both economically
inefficient and, as we shall see later, still less conservative than
the truly optimal policy for large stocks.   Other studies 
have likewise attempted to include uncertainties in model parameters and measurements
by making this same assumption that a constant escapement solution is optimal and 
thus avoiding solving the full sequential decision optimization problem [@Ludwig1981; @Moxnes2003; @Engen1997].

Sethi et al. [@Sethi2005] seek to address
the limitations of @Roughgarden1996 by searching for the true optimal
policy without the strong assumption of constant escapement.
Like @Clark1986 and others, @Sethi2005 finds support for the counter-intuitive result
of an optimal policy that is significantly less cautious when accounting
for measurement uncertainty than the corresponding constant escapement
solution which ignores it. 
Unfortunately, the approach taken by @Sethi2005 once again avoids solving
the full optimal decision-making problem by means of a simplification
which we might term the amnesic assumption: during each subsequent
decision the manager forgets any previous knowledge of the system.
This allows them to use the standard Stochastic Dynamic Programming
(SDP) solution method used by @Reed1979 and @Clark1986. This method
assumes that problem in question is a Markov Decision Process (MDP):
one in which given the state $x$ and action $a$, the probability of a
transition to state $x'$ is conditionally independent of all previous
states and actions (known as the Markov property).  This holds only if the state
is known without error -- once we introduce any amount of measurement
uncertainty about the state $x$, the process with respect to the observed
state is no longer Markovian but depends on all previous observations.

This problem can be solved by considering
the 'belief' state rather than the observed state, which restores
the Markov property at the cost of much greater problem complexity:
one must consider a (continuously-valued) probability
that the system is in each of the $|X|$ possible states of such a 
Partially Observed Markov Decision Process (POMDP). This is the
key feature of the POMDP approach: all previous observations are
collectively reflected in this belief distribution, and decisions
are based not only on the most recent observation (as in MDP), but
are also influenced by this prior belief.

The mechanics of 
updating these beliefs, while technical, have been well
illustrated elsewhere even in conservation applications
[e.g. @Lane1989; @Chades2008; @Haight2010; @Chades2011;
@Williams2011; @Springborn2013; @Fackler2014; @Fackler22014; @Baggio2016].
However, these previous studies have relied on relatively direct algorithms 
which are not amenable to the large number of states and actions required by the optimal
fisheries harvest problem.  By leveraging newer point-based algorithms developed in the artificial 
intelligence community [@Kurniawati2008; @Shani2013; see SI; R code], we are able to solve
the POMDP models for fisheries management directly. 
These solutions illustrate how incorporating the information
of all previous observations into a prior belief allows the
POMDP approach to handle measurement error which would 
cause a stock collapse under MDP or deterministic methods.
We also evaluate the performance of proposed method
on a real-world time series of population estimates of
marine ecosystems from RAM Legacy Stock
Assessment Database [@Ricard2011].

## Model formulation

Our formulation of the fisheries management problem closely follows that of 
earlier work [e.g. @Sethi2005], which we summarize here. Alternate formulations are provided in the SI,
illustrating that our conclusions are not constrained to these particular choices.

The fisheries manager seeks to maximize the expected net present value of the fishery
over an infinite time horizon under two sources of uncertainty: population dynamics (i.e. growth of the stock),
and stock measurements, given an model estimate of those dynamics:

\begin{equation}\label{optim}
\max_{\{h_t\} \geq 0} \mathbf{E} \left[ \sum_0^{\infty} \gamma^t U(x_t, h_t) \right]
\end{equation}

The state of the system, $x_t$ corresponds to the (true) fish stock at time $t$. 
The harvest is $h_t = \min \left(x_t,a_t\right)$, where $a_t$ is the attempted harvest at time step $t$.
For simplicity, we assume the reward (utility) is directly proportional to the
harvest, $U(x_t, h_t) = \min(h_t, x_t)$, with future rewards discounted by multiplicative discount factor $\gamma = 0.95$;
alternate harvest functions and alternate discount rate are considered in the SI.

The population dynamics is formulated as:

\begin{equation}
x_t = \epsilon_{g} f(s_{t-1})
\end{equation}

where $s_{t-1} = x_{t-1} - h_{t-1}$ is the escapement at time $t-1$, and
$\epsilon_{g}$ corresponds to stochastic component of recruitment (growth noise).

While it is theoretically possible to learn about this model and its parameters during the
decision process itself (so-called adaptive management, e.g. @Johnson2015 for MDP case; 
we discuss the potential for POMDP adaptive management in *Future Directions*), it is
typical to consider that state-space dynamics $f$ have been previously estimated from data.  
The goal of a decision process is to determine the optimal sequence of actions given this
model. From the population dynamics state
equation we derive a set of transition matrices $T(x_t, a_t, x_{t+1})$ describing the
probability of a transition from $x_t$ to $x_{t+1}$ under action $a_t$ (SI Methods).
While such transition matrices can be defined even when $f$ represents only a posterior
distribution of model parameter values [e.g. @Boettiger2015] or an age-structured 
model [e.g. @Holden2015], for simplicity of presentation we will focus on the case
in which $f$ is a fixed model with known parameters, given by the Ricker equation [@Ricker1954]:

\begin{equation}\label{ricker}
f(s_t) = s_t \exp\left(r \left(1-\frac{s_t}{K}\right)\right),
\end{equation}

where $K$ denotes the carrying capacity of the system, and $r$ the growth rate.
The optimal management of this model in the deterministic and growth-noise-only (MDP)
case is already well understood and thus will provide a useful foil against
which to examine the role of measurement uncertainty.  
To demonstrate that our results are not unique to this choice of \eqref{ricker}, two alternate models,
the Beverton-Holt model [@Beverton1957], and the Allen model [@Allen2005], are evaluated in the SI.

The POMDP problem is distinguished by the introduction of measurement error, in which 
the observed stock $z_t$ is measured with uncertainty, $z_t = \epsilon_{m} x_t$. 
In general this is defined by observation matrices $O(z_t, x_t, a_t)$ giving the 
probability of observing $z_t$ given a system in state $x_t$ and having taken action $a_t$,
though in our formulation this probability is independent of the action.

We will assume for simplicity that the stochastic growth noise $\epsilon_g$ and measurement error $\epsilon_m$ 
are distributed log-normally with mean $1$ and scale parameter $\sigma_{g}$ 
and $\sigma_m$ respectively. Larger $\sigma$ corresponds to higher variance,
scaled to be comparable to the uniform noise distribution shown in the SI,
(see *Materials and Methods*).  We evaluate the effect of
different noise levels in both growth and measurements, $\sigma_{g},
\sigma_{m} \in \{0, 0.1, 0.5\}$.

As the numerical solutions of the MDP and POMDP problems require discrete states,
we consider a system of 50 states on a uniform grid of 0 to 1 with $K = 0.66$ and
$r = 1$; (alternative growth rates shown in the SI).   
The transition matrices $T$, together with the observation matrix $O$ and the utility
function, completely specify a POMDP problem which can be solved with a numerical
algorithm (*Materials and Methods*).

# Results {#results .unnumbered}

To illustrate the impact of measurement error on management outcomes,
we compare two scenarios: with measurement error ($\sigma_m = 0.5$,
see *Materials and Methods*) and without measurement error (Fig \ref{sim}).
In each scenario, we compute the optimal policy using three different
solution methods: POMDP, MDP, and a policy we refer to as Offline POMDP.
In POMDP, the optimal policy at each time point depends not only on the
most recent observation, but on the manager's prior belief of the state,
which depends on all prior observations. As such, the optimal POMDP
policy for a particular observation cannot be computed ahead of time
but must be calculated "online", as observations come in.  In contrast,
in MDP, the optimal policy depends only on the most recent observation.
The "Offline" policy uses the POMDP solution given a fixed, uniform
prior belief, rather than updating the prior belief over states after
each observation.  For each scenario (with and without measurement error), 
we generate 100 simulations of  the fishery being managed under each 
strategy (MDP, Offline, POMDP) for 100 time steps.  We compute the average and standard
deviation of the stock size over these 100 replicates, Fig. \ref{sim},
as well as the mean economic return of each case (Table \ref{Table1})
and the fraction of simulated stocks that have collapsed by the end
of the simulation (Table \ref{Table2}). While we perform these
simulations over a range of models, parameter values and noise distributions, for
simplicity of presentation we focus on simulations
from a Ricker model with log-normally distributed noise (see *Materials and Methods*);
comparable results under the alternative choices are presented in the SI.

\begin{figure}[h]
\centering
\caption{Forward simulations of managing a marine ecosystem under POMDP and MDP models in the presence ( $\sigma_m = 0.5$, top panel) or absence ( $\sigma_m = 0$, bottom panel) of measurement error.  Lines show the mean stock size over 100 replicate simulations, error bands show +/- 1 standard deviation.  Many of the replicate simulations under MDP quickly collapse the stock by over-harvesting when measurement error is present.  In contrast, the POMDP management is robust to the level of measurement error, avoiding stock collapse whether or not measurement error is actually present.  The growth model is Ricker, the noise is log-normally distributed, the carrying capacity of the system is $K = 0.66$, the maximum per capita growth rate is $r = 1$, growth noise is $\sigma_g = 0.5$, and the discount factor is $\gamma = 0.95$.}\label{sim}
\end{figure}

Fig. \ref{sim}, top panel, illustrates that ignoring or underestimating
the measurement error result in dramatic collapse of the stock and
significant loss in revenue: the mean MDP state declines as subsequent
replicate simulations collapse (94/100 collapse within 100 time steps)
when measurement error leads to over-harvesting.  The POMDP method
retains a steady, sustainable harvest that keeps the population below
carrying capacity but results in minimal population collapse (4/100). The
offline solution results in some stock collapse (34/100) as a result of
not updating the manager's belief state over subsequent observations.
Unsurprisingly, average economic returns are also much lower for MDP
when measurement error is present, returning an average of 66% of the net present value returned by
POMDP (Table \ref{Table1}), and would be lower still with less aggressive discounting of future
rewards (Here $\gamma = 0.95$ makes later collapses less relevant to the total).
Fig. S1 shows stock trajectories from several individual replicates, which 
show relatively little decline before collapsing suddenly when managed under MDP
when measurement error is present.  Averaging over an ensemble of such trajectories
gives the appearance of a smooth decline.

On the other hand, overestimating measurement error might result in a
trivial loss in revenue but does not result in collapse of the stock,
(Fig. \ref{sim}, lower panel). In this scenario, since no
measurement error is present, the MDP solution is indeed optimal, but the
POMDP policy which assumes a measurement error of $\sigma_m = 0.5$ and
thus overestimates the measurement uncertainty performs nearly as well
(average of 94% of the MDP net present value and likewise no collapses of the stock, Table \ref{Table2}).
In the SI we report the same patterns for Ricker model
with uniform noise (Fig. S2), Beverton-Holt model with log-normal noise
(Fig. S3), and Allen model with log-normal noise (Fig. S4).

<!-- Perhaps in-text quotations of these numbers is sufficient?-->

\begin{table}
\centering
\caption{Ecological and economic performance when
measurement error is present, (\(\sigma_m = 0.5\))}
\begin{tabular}{lccc}
& MDP & Offline & POMDP \\
\midrule
Fraction collapsed & 94\% & 34\% & 4\% \\
Economic return & 66\% & 99\% & 100\% \\
\bottomrule
\end{tabular}
\label{Table1}
\end{table}

\begin{table}
\centering
\caption{Ecological and economic performance when
measurement are perfect, (\(\sigma_m = 0\))}
\begin{tabular}{lccc}
& MDP & Offline & POMDP \\
\midrule
Fraction collapsed & 0\% & 0\% & 0\% \\
Economic return & 100\% & 96\% & 94\% \\
\bottomrule
\end{tabular}
\label{Table2}
\end{table}

To facilitate comparison with previous research and management practice,
we plot the corresponding optimal action in terms of *expected escapement*: the
stock the manager expects will be left in the sea after harvest $z_t - h_t$.
Recall that introducing measurement error means that this value can differ
from realized escapement, $x_t - h_t$. Fig. \ref{policy} shows the optimal expected 
escapement as a function of observed stock sizes over a range of growth 
noise and observation error intensities, assuming a uniform prior belief over
states $x_t \in (0,1)$ for $\sigma_m > 0$. These policies are determined numerically
using the SARSOP algorithm for POMDP [ @Kurniawati2008], as discussed in *Materials and Methods* and
illustrated in the supplementary code provided.

In the MDP case ($\sigma_m = 0$),
we see the classical result of constant escapement [@Reed1979], where the stock is harvested
back down to the same escapement regardless of observed stock or growth noise.
When growth noise is large ($\sigma_g = 0.5$) relative to measurement
error ($\sigma_m = 0.1$), the policy converges towards the the constant
optimal escapement solution of @Reed1979, as expected. These patterns
hold independent of the population growth model (e.g. Beverton-Holt and
Allen) and the shape of the noise distribution (Fig. S5--S6).
The most obvious difference between the MDP and POMDP policies arises in the region of
linear or nearly linear increases in expected escapement occurring at very large observed stock sizes,
which arises from any policy whose harvest remains *constant* over that range. Not increasing
harvest in response to higher observed stock sizes always result in a higher expected escapement, $z_t - h_t$.  
This corresponds to a policy which believes such high observations are almost surely the 
result of measurement error, and suggests that overestimated stock sizes are responsible for MDP
over-harvesting and the subsequent collapse of the stock. At smaller stock sizes, the policies
shown in Fig. \ref{policy} appear more similar over any measurement error, but in fact this is only a consequence of the
uniform prior belief. While Fig. \ref{policy} completely specifies the management policy under MDP,
this is not true for the POMDP solutions ($\sigma_m > 0$), which also depend on all
prior observations through the influence of the belief state.

\begin{figure}[h]
\centering
\caption{Optimal escapement policy as a function of measured stock for Ricker model, with different measurement errors and growth noise. Noises are log-normally distributed, and the prior belief state is uniform among all stocks. The carrying capacity of the system is $K = 0.66$, the growth rate is $r = 1$ and the discount factor is $\gamma = 0.95$.}\label{policy}
\end{figure}

To illustrate this key difference, we plot (Fig. \ref{bel})
the POMDP policy under two alternative prior belief states: (1) Low
Stock: where the prior belief is centered at $x = 0.2$ , and (2) High Stock:
where the prior belief is centered at the location of carrying capacity $x =
0.66$. This figure is generated in an identical manner to Fig. \ref{policy} 
except for the different choice of prior belief.  Visualizations of these 
prior distributions are shown in Fig. S12. When the prior belief skews towards low stocks, the
probability that an observation of a high stock is an overestimate due to
measurement error is greater, and consequently the policy becomes more conservative
(higher escapement). This ability to incorporate the knowledge from prior
observations accounts for the difference
in performance between the POMDP solution and the Offline solution in
Fig. \ref{sim} when measurement error is present. 
The evolution of the belief state in response to subsequent observations
over a given simulation is shown for the different models in Figs S14 - S16.
Corresponding plots for alternate models are shown in Fig. S9.
The patterns shown here are further robust to the reward function (including
linear cost to harvest and a diminishing returns model, Fig. S8) and the
discount rate Fig. S13.

\begin{figure}[h]
\centering
\caption{Optimal escapement policy as a function of measured stock and different prior belief states: (1) Uniform: uniform prior belief over all stocks, (2) Low Stock: the prior belief is centered at $x = 0.2$, and (3) High Stock: the prior belief is centered at the location of carrying capacity $x =
0.66$. In this figure, growth model is Ricker, measurement and growth noises are $\sigma_m = \sigma_g = 0.5$, and are log-normally distributed, the growth rate is $r = 1$, and the discount factor is $\gamma = 0.95$.}\label{bel}
\end{figure}

## Application to real-world fisheries

>>>>

Imperfect information is ubiquitous in ecological management and conservation decision making. While the pressing concerns of global change have put the spotlight on *forecasting*, i.e. the uncertainty of the future [e.g. @Petchy2015], management decisions must also contend with uncertainty of the *present*: How many fish are in the sea today? What regions harbor the most fragile biodiversity? How have invasive species transformed native communities? A recognition of the importance of uncertainty as well as sophisticated methods and tools to do so has become increasingly widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierachical Bayesian Modeling, along with convenient software tools for implementing these methods.  In sharp contrast to that shift, the treatment of uncertainty in decisions -- how the uncertainty in models should be reflected in scientific recommendations, remains far less developed, more qualitative, and underserved by available tools.  In this paper we illustrate how uncertainty in measurement has a dramatic negative impact on outcomes under a variety of these exsiting approaches, which can be avoided through the use of more formal approaches that have since been developed in other fields that can likewise be implemented in software tools.  This insight both resolves a long-standing paradox in fisheries management and has important implications for how ecosystem management deals with uncertainty more generally.

@@

The area of fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1957], thanks to its global relevance, long history, and readily available data.  Consequently, understanding the impact of measurement uncertainty in this rich and well-studied context will have implications for both resource management and conservation efforts in other domains, many of which draw on concepts established or tested in fisheries ecosystems.  We compare the results of managing simulated fish stocks under the existing optimal control approach, a heuristic approach that more closely resembles today's management practices, and a newly proposed approach that provides a more rigorous treatment of uncertainty.

@@

Numerically, each of these functions are defined over a discrete set of possible states, observations, and actions, and can thus be represented as a collection of matrices or tensors.  As in the case of MDP [@Marescot2013], the POMDP problem can be solved by existing and well-studied algorithms. While not unknown to the conservation literature [@Williams2009], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g. @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b].  POMDP algorithms remain a very active area in artificial intellegence, and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries managment, which requires on the order of 100 states and actions to provide sufficient numerical resolution. Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods. For those interested in exploring or applying these methods, we have developed a preliminary R package, @sarsop, which can be used to replicate any of the results presented here and explore other variations.

@@

```{r}
```

@@
